AAAI Press Formatting Instructions \\for Authors Using \LaTeX{} --- A Guide

### 2021-09-17 publications/iaai21/dora_outlier_detection_iaai.tex
Minor updates to text, formatting, todos. #42

\section{Experiments}

\subsection{Datasets}
We constructed three datasets to evaluate the utility of DORA and the
performance of each ranking algorithm for a variety of scientific domains
 (astrophysics, planetary science, and Earth science). We also included a 
 benchmark dataset that uses MNIST and Fashion-MNIST. Here we describe
 the details of each dataset.
\todo{Maybe add
 a table with number of samples, number of reference known outliers (if 
 available), etc, if room? }
 \todo{Can we make any of these datasets publicly available with the paper?}

@@

\subsubsection{MNIST and Fashion-MNIST.}
We included MNIST and F-MNIST to demonstrate the capabilities of DORA with a
traditional benchmark dataset. In this experiment, we used 60,000 images from 
F-MNIST as the training set and a combined set of 2,000 images from MNIST and 
F-MNIST as the evaluation set. Well-discriminating algorithms should assign 
higher outlier scores to samples from MNIST than F-MNIST.

@@

    \hline
    Astrophysics &   &   &   \\
    Earth & 6,757 & 37 & 76  \\
    Planetary &   &   &   \\
    F-MNIST & 60,000 & 1,000 & 1,000  \\
    \hline
  \end{tabular}
\end{table}

\subsection{Results}
The experimental setup for each dataset was to fit or train a model for
each ranking algorithm using a larger unlabeled dataset and then apply 
the models to compute the outlier scores for a smaller test dataset for which
labels of known outliers were available (Table \ref{tab:datasets}). 
For each test set, we created a plot of the number of known/true outliers 
detected out of the top $N$ selections. We also 
reported the Mean Discovery Rate (MDR) in the legend
for each algorithm to give a quantitative comparison of the curves that could
be comparable across the datasets. We defined MDR as:
\begin{equation}
MDR = \frac{\sum^{N_s}_{i=1} n_i}{\sum^{N_s}_{i=1} s_i}
\end{equation}
where $i \in [1, N_s]$ is the selection index,
 $N_s$ is the total number of selections,
 $s_i$ is the number of selections made up to index $i$,
 and $n_i$ is the number of known outliers (true positives)
among $s_i$ selections.
We also reported the precision at $N=n_{outlier}$ for each test set where
$n_{outlier}$ is the number of known outliers, i.e.,  the precision obtained 
when the number of selections is the same as the total number of outliers. 
Precision at $N$ is the number of true outliers divided by the number of 
selections $N$~\citep{campos2016evaluation}.
Table~\ref{tab:faw_results} compares the precision at $N=n_{outlier}$ for
each dataset and ranking algorithm. 
%\todo{Hannah: Consider merging all three data set discussions since they are
%in one table? with their own figures? The point here isn't really to
%compare algs or data sets but instead to show that we CAN compare
%them, using DORA.}
% We initialized each algorithm with $D_s$ and requested a ranking of
% targets in $D_{s+}$.
%

@@

For the MNIST and F-MNIST dataset, PCA and DEMUD tied for the highest precision,
while DEMUD, PCA, and PAE tied for the highest MDR. Negative sampling had the 
lowest performance in both metrics.

\begin{table}
  \caption{Precision at $N=n_{outlier}$ for three applications; the
  best result for each data set is in bold.}
  \label{tab:faw_results}
  \centering
  \begin{tabular}{l|cccc}
    \hline
    Algorithm & Astro. & Earth & Planetary & F-MNIST \\
    \hline
    PCA               & & 0.41 & 0.44 & \textbf{0.84} \\
    DEMUD             & & 0.41 & 0.44 & \textbf{0.84} \\
    \hline
    RX                & & 0.43 & ---- & 0.82 \\
    LRX               & & ---- & 0.33 & 0.56 \\
    \hline
    Iso. Forest  & & 0.46 & \textbf{0.56} & 0.74 \\
    \hline
    PAE               & & 0.41 & 0.44 & 0.83 \\
    Neg. Sampling & & \textbf{0.49} & 0.33 & 0.43 \\
    \hline
    Random            & & 0.32 & 0.11 & 0.49 \\
    \hline
  \end{tabular}
\end{table}

@@

\section{Discussion}
\todo{ 
Potential topics TBD based on results:
\begin{itemize}
\item Discuss results from experiments
\item Challenges of evaluating results in outlier detection?
\item Downstream use of results in science applications
\end{itemize}
}
\paragraph{Path to deployment.} \todo{Hannah - describe pip-installable
package, open-source for extension by the wider scientific community, etc.}

>>>>

\section{Datasets}

We constructed three datasets to evaluate the utility of DORA and the
performance of each ranking algorithm for a variety of scientific domains
 (astrophysics, planetary science, and Earth science). We also included a 
 benchmark dataset that uses MNIST and Fashion-MNIST. Here we describe
 the details of each dataset.
\todo{Maybe add
 a table with number of samples, number of reference known outliers (if 
 available), etc, if room? }
 \todo{Can we make any of these datasets publicly available with the paper?}

@@

\subsubsection{MNIST and Fashion-MNIST.}
We included MNIST and F-MNIST to demonstrate the capabilities of DORA with a
traditional benchmark dataset. We used 60,000 images from 
F-MNIST as the training set and a combined set of 2,000 images from MNIST and 
F-MNIST as the evaluation set. Well-discriminating algorithms should assign 
higher outlier scores to samples from MNIST than F-MNIST.

@@

    \hline
    Astrophysics &   &   &   \\
    Earth & 6,757 & 37 & 76  \\
    Planetary & 992 &  9 & 49 \\
    F-MNIST & 60,000 & 1,000 & 1,000  \\
    \hline
  \end{tabular}
\end{table}

\section{Results}
The experimental setup for each dataset was to fit or train a model for
each ranking algorithm using a larger unlabeled dataset and then apply 
the models to compute the outlier scores for a smaller test dataset for which
labels of known outliers were available (Table \ref{tab:datasets}). 
For each test set, we created a plot of the number of known/true outliers 
detected out of the top $N$ selections. We also 
reported the Mean Discovery Rate (MDR) in the legend
for each algorithm to give a quantitative comparison of the curves that could
be comparable across the datasets. We defined MDR as:
\begin{equation}
MDR = \frac{\sum^{N_s}_{i=1} n_i}{\sum^{N_s}_{i=1} s_i}
\end{equation}
where $i \in [1, N_s]$ is the selection index,
 $N_s$ is the total number of selections,
 $s_i$ is the number of selections made up to index $i$,
 and $n_i$ is the number of known outliers (true positives)
among $s_i$ selections.
We also reported the precision at $N=n_{outlier}$ for each test set where
$n_{outlier}$ is the number of known outliers, i.e.,  the precision obtained 
when the number of selections is the same as the total number of outliers. 
Precision at $N$ is the number of true outliers divided by the number of 
selections $N$~\citep{campos2016evaluation}.
Table~\ref{tab:faw_results} compares the precision at $N=n_{outlier}$ for
each dataset and ranking algorithm. 
\todo{Jake: add description of theoretical random baseline.}
%\todo{Hannah: Consider merging all three data set discussions since they are
%in one table? with their own figures? The point here isn't really to
%compare algs or data sets but instead to show that we CAN compare
%them, using DORA.}
% We initialized each algorithm with $D_s$ and requested a ranking of
% targets in $D_{s+}$.
%

@@

For the MNIST and F-MNIST dataset, PCA and DEMUD tied for the highest
 precision at $N=n_{outlier}$
while DEMUD, PCA, and PAE tied for the highest MDR. Negative sampling had the 
lowest performance in both metrics.

\begin{table}
  \caption{Precision at $N=n_{outlier}$ for three applications and F-MNIST
   benchmark; the best result for each data set is in bold.}
  \label{tab:faw_results}
  \centering
  \begin{tabular}{l|cccc}
    \hline
    Algorithm & Astro. & Earth & Planetary & F-MNIST \\
    \hline
    PCA               & & 0.41 & 0.44 & \textbf{0.84} \\
    DEMUD             & & 0.41 & 0.44 & \textbf{0.84} \\
    \hline
    RX                & & 0.43 & ---- & 0.82 \\
    LRX               & & ---- & 0.33 & 0.56 \\
    \hline
    Iso. Forest  & & 0.46 & \textbf{0.56} & 0.74 \\
    \hline
    PAE               & & 0.41 & 0.44 & 0.83 \\
    Neg. Sampling & & \textbf{0.49} & 0.33 & 0.43 \\
    \hline
    Random            & & 0.32 & 0.11 & 0.49 \\
    \hline
  \end{tabular}
\end{table}

@@

\section{Discussion}

\paragraph{Algorithm performance.} \todo{Discuss overall results from
 experiments and that we left hyperparameter tuning and extensive
 evaluation for future work.}

\paragraph{Evaluation in outlier detection.} \todo{Discuss why evaluation
in outlier detection applications is hard since we often don't have complete
labeled data for testing and labels can be subjective.}

\paragraph{Path to deployment.} \todo{Hannah - describe pip-installable
package, open-source for extension by the wider scientific community, etc.}
